On the Uncertainty of Wind Power Forecasts

### 2019-08-28 LaTeX_code/paper/paper_v4.tex
updated tripizoidal

%%%%%%%%% TITLE
\title{Paper Plan  \\ "Stochastic Wind Power Forecasting"  }  % \\  \small{Report}}

@@

\begin{itemize}
    \item Briefly say the motivation and applications
    \item What is novel here and why its better.
    \item summarize the technique
    \item mention the data used.
    \item mention concluding results.
\end{itemize}

@@

\begin{itemize}
    \item Importance and interested parties
    \item Literature review on current forecasting practices
    \item shortcomings of current deterministic and probabilistic forecasts.
    \item Mention our ability of simulating production and why it's important.
    \item Consequence of not taking into account the real world performance of the forecasts. And present proofs where such forecasts fail such as in complex terrains.
    \item Consequences of not taking into account skewness of the errors this includes asymmetric cost for up and down-regulation of power.
    \item Emphasize that it works independent of the forecasting technology. Thus we are able to compare different forecasts based on their real world performance.
    \item Mention past contributions to the topic and show what is novel in this approach.
    \item Mention the data used in this study.
\end{itemize}

@@

\begin{itemize}
    \item Introduce why we have the physical constrains.
    \item Discuss the normalization between $[0,1]$ and why its needed
\end{itemize}

@@

\begin{itemize}
    \item combine both drift and diffusion control
    \item Do a Feller test to show that the process stays almost surely in $[0,1]$
    \item Show that if we start with zero error it will stay that way indefinitely.

@@

\begin{itemize}
    \item say that we have two main approaches based on moment matching:
    \begin{itemize}
        \item state dependent diffusion and use either a  Beta distribution as a proxy or a maximum entropy distribution.
        \item state-independent diffusion and use the gaussian as a proxy either by linearizing the SDE or its moments equations.
    \end{itemize}
    \item introduce Lamperti transformation on $V$, write the new SDE and its moments.
    \item give explicit expression to compute the moments as a proposition.
    \item Say that We may infer in either spaces, the original space or Lamperti space.
    \item Discuss why we approximate the SDE directly by linearization or why to approximate its moments instead.

@@

\section{Parameter Estimation}
\begin{itemize}
    \item Write the likelihood for one path first using markovianity.
    \item Justify the product of paths by introducing observation error to the recoded production observations. Write that it's independent and small Gaussian noise.
    \item Mention that the Fokker-Plank written in the model section is computationally formidable for parameter estimation.
    \item Mention the two approaches: State dependent diffusion and state-independent diffusion SDE.
    \item discuss that state dependent diffusion SDE approach assumes a proxy distribution either a Beta or a maximum entropy distribution. That this is the underlaying assumption and reason it based on the skewness and boundedness of the data explained in the above section.
    \item discuss that the state-independent approach requires either linearizing the SDE or its moments equations. Justify it by saying that the time intervals are small and such approximations are reasonable.
\end{itemize}

@@

\section{Results}
For each model and technique we will state the following as a results:
\begin{itemize}
    \item Results from optimization
    \begin{itemize}
        \item Convergence of the ellipse around the optimal point.
        \item confidence intervals on obtained the parameters
        \item AIC and BIC information criteria.
    \end{itemize}
    \item Plots of forecast simulations.
    \item Plots of forecast confidence intervals.
\end{itemize}

@@

\section{Model on Disaggregated Data}
Choose the best forecasting company, model, technique and  apply it on the disaggregated data. State results as before for each wind farm.

\section{Extra: Optimal forecasting update interval}
\begin{itemize}
    \item We have hourly updated forecast from France. Therefore we can choose the best series of predictions (the first point of each update which is at an hourly frequency).
    \item  Allows us to evaluate the uncertainty for the "next hour" wind power production forecasts.
    \item redo the same evaluation on series made up of updates every 4 hours and another every 6 hours and so on.
    \item this allows us to find the optimal update frequency for different operational uses and clients.
    \item we are able to classify forecasts according to which forecast horizon they are better at. Say forecaster A technology is better at forecasting the next 12 hours and must be recomputed at that frequency. However, forecaster B is better at forecasting the next hour and must be recomputed hourly.
\end{itemize}

>>>>

%%%%%%%%% TITLE
\title{Paper Outline  \\ "Stochastic Wind Power Forecasting"  }  % \\  \small{Report}}

@@

\begin{itemize}
    \item Briefly say the motivation and applications
    \item What is novel here and why its better.
    \begin{itemize}
        \item Fast
        \item accurate
        \item computationally efficient
    \end{itemize}
    \item summarize the technique
    \begin{itemize}
        \item Indirect inference
        \item Moment based approximation of the transition densities
    \end{itemize}
    \item mention the data used.
    \begin{itemize}
        \item Uruguay
        \item frequency
        \item Period
    \end{itemize}
    \item mention concluding results.
\end{itemize}

@@

\begin{itemize}
    \item Importance and interested parties
    \item Literature review on current forecasting practices
    \item shortcomings of current deterministic and probabilistic forecasts.
    \item Mention our ability of simulating production and why it's important.
    \item Consequence of not taking into account the real world performance of the forecasts. And present proofs where such forecasts fail such as in complex terrains.
    \item Consequences of not taking into account skewness of the errors this includes asymmetric cost for up and down-regulation of power.
    \item Emphasize that it works independent of the forecasting technology. Thus we are able to compare different forecasts based on their real world performance.
    \item Mention past contributions to the topic and show what is novel in this approach.
    \item Mention the data used in this study.
    \item Summary of the structure of the paper, describe the flow and what each section contains.
\end{itemize}

@@

\begin{itemize}
    \item Introduce why we have the physical constrains.
    \item We want to fulfill the physical constraints with the least number of parameters and assumptions (parsimony).
    \item Discuss the normalization between $[0,1]$ and why its needed
\end{itemize}

@@

\begin{itemize}
    \item write the model including the drift and diffusion.
    \item Do a Feller test to show that the process stays almost surely in $[0,1]$
    \item Show that if we start with zero error in the mean it will stay that way indefinitely. Show this using Ito.

@@

\begin{itemize}
    \item note that the state-independent diffusion model is favorable to utilize probability theory.
    \item introduce Lamperti transformation on $V$, write the new SDE and its moments.
    \item give explicit expression to compute the moments as a proposition.
    \item Remark that in the lamperti space:
    \begin{itemize}
        \item no longer able to obtain the exact moment equations as we had in the original space.
        \item inexactness of the moment equations introduces a bias which we didn't have in the original space.
    \end{itemize}
    \item We can work in either the original space (state dependent diffusion) or the Lamperti space (state-independent diffusion).
    \begin{itemize}
        \item In the original space we use either a  Beta distribution as a proxy or a maximum entropy distribution.
        \item In the Lamperti space we use the gaussian as a proxy either by linearizing the SDE or its moments equations.
    \end{itemize}
    \item Say that We may infer in either spaces, the original space or Lamperti space.
    \item Discuss why we approximate the SDE directly by linearization or why to approximate its moments instead.
    \item Ability to estimate parameters using the brackets and least squares when high-frequency data is available.

@@

\section{Parameter Estimation}
\begin{itemize}
    \item Write the likelihood for one path first using markovianity.
    \item Justify the product of paths by introducing observation error to the recoded production observations. Write that it's independent and small Gaussian noise.
    \item Mention that the Fokker-Plank written in the model section is computationally formidable for parameter estimation.

    \item Mention the two spaces: the original space and Lamperti space.

    \item We assumes a proxy distribution either a Beta or a maximum entropy distribution. That this is the underlaying assumption and reason it based on the skewness and boundedness of the data explained in the above section.
    \item In Lamperti space discuss that we either linearize the SDE or its moments equations. Justify it by saying that the time intervals are small and such approximations are reasonable.
\end{itemize}

@@

\section{Results}
For each model and technique we will state the following as a results:
\begin{itemize}
    \item Results from optimization
    \begin{itemize}
        \item Convergence of the ellipse around the optimal point.
        \item confidence intervals on obtained the parameters
        \item AIC and BIC information criteria for model comparison.
    \end{itemize}
    \item Results from MCMC to verify and demonstrate that we have exhaustively searched the space. This is important as we are not able to guarantee the convexity of our objective function.
    \item Plots of forecast simulations.
    \item Plots of forecast confidence intervals.
\end{itemize}

@@

\begin{itemize}
    \item Assigning weights for making forecast ensembles.
\end{itemize}

\section{Model on Disaggregated Data}
Choose the best forecasting company, model, technique and  apply it on the disaggregated data. State results as before for each wind farm.
\begin{itemize}
    \item We study it as a network and how it behaves for power transport and load management. This means that we cannot assume that the nodes are independent, but we need to capture the interplay between them with a minimal number of parameters to tune.
    \item Must incorporate correlation between wind farms in the noise $dW_t$.
\end{itemize}
\section{Extra: Optimal forecasting update interval}
\begin{itemize}
    \item We have hourly updated forecast from France. Therefore we can choose the best series of predictions (the first point of each update which is at an hourly frequency).
    \item  Allows us to evaluate the uncertainty for the "next hour" wind power production forecasts.
    \item redo the same evaluation on series made up of updates every 4 hours and another every 6 hours and so on.
    \item this allows us to find the optimal update frequency for different operational uses and clients.
    \item we are able to classify forecasts according to which forecast horizon they are better at. Say forecaster A technology is better at forecasting the next 12 hours and must be recomputed at that frequency. However, forecaster B is better at forecasting the next hour and must be recomputed hourly.
\end{itemize}

### 2019-08-28 LaTeX_code/poster/poster.tex
p(1-p) removed,added path simulator,path length feature, non-tech poster
updated tripizoidal

@@

\subsection*{Physical Constrains}
Let $p_t$ be a numerical wind power forecast, which is an input to this approach. Then the model is given by the following It\^{o} stochastic differential equation,
\begin{equation}
\begin{split}
dX_t&= \dot{p} \ dt + (\theta_t -X_t) \ dt + b (X_t; \bm{\theta} ) \ dW_t \quad t > 0 \\
X_0&=x_0
\end{split}
\end{equation}
As the  installed power production capacity is limited, we normalize with respect to the installation power. Thus our process must be limited to the range $[0,1]$. To enforce this constraint, we must have drift and diffusion control.
\subsubsection*{Diffusion Control: } The physical constraint is respected  by choosing  diffusion coefficient $ b (x; \bm{\theta} )= \sqrt{2 \theta_t \alpha p_t(1-p_t) x (1-x)} $ that is zero on the boundaries of $[0,1]$.
\subsubsection*{Drift Control: }
$X_t$ remains in $[0,1]$ a.s. by choosing $\theta_t$ as follows, 
\begin{equation}
\theta_t = \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )
\end{equation}

\subsubsection*{Change of Variables:}
In order to avoid differentiation of the forecast $\dot{p}_t$ and simplify, we apply a change of variables $$V_t = X_t - p_t$$  \\
Then  model becomes,
\begin{equation}
\begin{split}
dV_t &=  - \theta V_t \  dt + \sqrt{2 \theta \alpha p_t(1-p_t) (V_t +p_t ) (1-V_t-p_t)} \  dW_t \quad t > 0 \\
V_0 & = v_0
\end{split}\label{main}
\end{equation}

@@

We apply the model to Uruguayan wind power generation data and their corresponding numerical wind power production forecast. \\

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{Forecast_data_68.pdf}
    \includegraphics[width=0.8\linewidth]{Forecast_data_82.pdf}
\end{center}
   \caption{Examples of wind power generation from the data set along with the wind power generation forecast.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

@@

\subsection*{ Optimization  }
\begin{itemize}
%\item The log-likelihood is convex in the  parameters $(\alpha, \theta_t )$ on the positive quadrant.
\item We optimize using L-BFGS algorithm constrained to the positive quadrant.
\end{itemize}
The ellipse defined by the  Hessian of the log-likelihood shrinks at a fast rate as shown in Figure (\ref{ellipse_drawing}). The contours of the  log-likelihood can be seen in Figure (\ref{contour}).

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.6\linewidth]{ellipse1193_samples_dN=14e-02.pdf}
    \includegraphics[width=0.6\linewidth]{ellipse_conv_samples_dN=14e-02.pdf}
\end{center}
   \caption{Left:Shrinkage of the ellipse determined by the Hessian matrix of the log-likelihood around the point of optimality $(\theta_0^*, \alpha^*)\approx (8,1)$ (Axis are not in natural scale, arrows only to  eignevector direction). Right: Convergence of the major and semi-axis of the ellipse  of the Hessian of the log-likelihood at the point optimality $(\theta_0^*, \alpha^*)\approx (8,1)$. Note that it is slightly faster than the expected rate of $1/\sqrt{M}$. This is due to the correlation structure of the process $V_t$, thus a path may act  as more than one  uncorrelated sample.}
\label{ellipse_drawing}
\end{figure}

 \begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.9\linewidth]{ISO_100_samples_dN=14e-02.pdf}
   %\includegraphics[width=0.5\linewidth]{ISO_500_samples_dN=14e-02.pdf}
   %\includegraphics[width=0.5\linewidth]{ISO_1193_samples_dN=14e-02.pdf}
\end{center}
   \caption{Contour plot of the log-likelihood with only 100 sample paths, point of optimality $(\theta_0^*, \alpha^*)\approx (8,1)$ indicated by the  black dot.}
\label{contour}
\end{figure}

\section*{Results}
 We are able to obtain the parameters based on the complete data sets mentioned earlier. The parameters are given by $(\theta^*, \alpha^*)\approx (8,1)$.

%We can also see in Figure() that the region surrounding the optimal parameters becomes more defined as we add more sample paths.

In Figure (\ref{fig:72hr}) and (\ref{fig:6hr}), we simulate sample forecast errors using the optimal parameters  $(\theta_0^*, \alpha^*)\approx (8,1)$ and obtain the associated confidence bands empirically using 1000 sample paths for each forecast. %Note that the confidence bands obtained are non-trivial and reflect the complexity of the uncertainties in numerical wind power generation forecasts.

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{72hr_forecast_CI_31.pdf} %667
   \includegraphics[width=0.8\linewidth]{72hr_forecast_CI_437.pdf}
   \includegraphics[width=0.8\linewidth]{72hr_forecast_CI_82.pdf}

\end{center}
   \caption{ Examples of confidence bands obtained for the full 72 hour forecasts.We can see that the model captures the fluctuations in the actual production with non-trivial and asymmetric confidence intervals.}
\label{fig:72hr}
\end{figure}

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{6hr_forecast_CI_75.pdf}  %569
   \includegraphics[width=0.8\linewidth]{6hr_forecast_CI_59.pdf}
   %\includegraphics[width=0.9\linewidth]{6hr_forecast_CI_68.pdf}
\end{center}
   \caption{ Examples of confidence bands obtained for the first 6 hours of the forecasts. This is important as this specific forecasting company computes a new forecast every 6 hours for reliability.}
\label{fig:6hr}
\end{figure}

@@

\section*{Conclusions}
In this project, we have proposed a model based on SDE's to quantify uncertainties in wind power generation forecasts. It has the following advantages:
\begin{itemize}
\item  Represents and quantifies  uncertainty in wind power forecasts.
\item It is forecast technology agnostic.
\item  Takes into account physical constrains and the skew-symmetric nature of forecast error.
\item Provides a basis for decision making in the optimal dispatch of electric power.
\end{itemize}

@@

\section*{References}
\begin{flushleft}
[1] {Elkantassi, S., Kalligiannaki, E., \& Tempone, R. (2017). Inference And Sensitivity In Stochastic Wind Power Forecast Models. Proceedings of the 2nd International Conference on Uncertainty Quantification in Computational Sciences and Engineering (UNCECOMP 2017). doi:10.7712/120217.5377.16899}
\end{flushleft}

>>>>

To capture real-world performance of  wind power forecasts, our model incorporates historical wind power production data along with their corresponding forecasts.\\

We apply the model to Uruguayan wind power generation data and their corresponding numerical wind power production forecasts. \\

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{Forecast_data_68.pdf}
    \includegraphics[width=0.8\linewidth]{Forecast_data_82.pdf}
\end{center}
   \caption{Examples of wind power generation from the data set along with the wind power generation forecast.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

@@

\subsection*{Physical Constrains}
Let $p_t$ be a numerical wind power forecast, which is an input to this approach. Then the model is given by the following It\^{o} stochastic differential equation,
\begin{equation}
\begin{split}
dX_t&= \dot{p} \ dt - \theta_t( X_t- p_t) \ dt + b (X_t; \bm{\theta} ) \ dW_t \quad t > 0 \\
X_0&=x_0
\end{split}
\end{equation}
As the  installed power production capacity is limited, we normalize with respect to the installation power. Thus our process must be limited to the range $[0,1]$. To enforce this constraint, we must have drift and diffusion control.
\subsubsection*{Diffusion: } The physical constraint is respected  by choosing  diffusion coefficient $ b (x; \bm{\theta} )= \sqrt{2 \theta_t \alpha x (1-x)} $ that is zero on the boundaries of $[0,1]$.
\subsubsection*{Drift: }
$X_t$ remains in $[0,1]$ a.s. by choosing $\theta_t$ as follows,
\begin{equation}
\theta_t = \max \left( \theta_0 \ , \ \frac{|\dot{p}_t|}{\min (p_t, 1-p_t)}  \right )
\end{equation}

\subsubsection*{Change of Variables:}
In order to avoid differentiation of the forecast $\dot{p}_t$ and simplify, we apply a change of variables $V_t = X_t - p_t$. Then  model becomes,
\begin{equation}
\begin{split}
dV_t &=  - \theta_t V_t \  dt + \sqrt{2 \theta_t \alpha (V_t +p_t ) (1-V_t-p_t)} \  dW_t \quad t > 0 \\
V_0 & = v_0
\end{split}\label{main}
\end{equation}

@@

@@

\subsection*{ Optimization  }
 \begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.9\linewidth]{ISO_100_samples_dN=14e-02.pdf}
   %\includegraphics[width=0.5\linewidth]{ISO_500_samples_dN=14e-02.pdf}
   %\includegraphics[width=0.5\linewidth]{ISO_1193_samples_dN=14e-02.pdf}
\end{center}
   \caption{Contour plot of the log-likelihood with only 100 sample paths, point of optimality $(\theta_0^*, \alpha^*)\approx (8,1)$ indicated by the  black dot.}
\label{contour}
\end{figure}
\begin{itemize}
%\item The log-likelihood is convex in the  parameters $(\alpha, \theta_t )$ on the positive quadrant.
\item We optimize using L-BFGS algorithm constrained to the positive quadrant.
\end{itemize}
The ellipse defined by the  Hessian of the log-likelihood shrinks at a fast rate as shown in Figure (\ref{ellipse_drawing}). The contours of the  log-likelihood can be seen in Figure (\ref{contour}).

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    \includegraphics[width=0.6\linewidth]{ellipse_conv_samples_dN=14e-02.pdf}
   \includegraphics[width=0.6\linewidth]{ellipse1193_samples_dN=14e-02.pdf}
\end{center}
   \caption{Left:Shrinkage of the ellipse determined by the Hessian matrix of the log-likelihood around the point of optimality $(\theta_0^*, \alpha^*)\approx (8,1)$ (Axis are not in natural scale, arrows only to  eignevector direction). Right: Convergence of the major and semi-axis of the ellipse  of the Hessian of the log-likelihood at the point optimality $(\theta_0^*, \alpha^*)\approx (8,1)$. Note that it is slightly faster than the expected rate of $1/\sqrt{M}$. This is due to the correlation structure of the process $V_t$, thus a path may act  as more than one  uncorrelated sample.}
\label{ellipse_drawing}
\end{figure}

\section*{Results}
In the following figures , we simulate sample wind power production forecasts using the optimal parameters  $(\theta_0^*, \alpha^*)\approx (8,1)$ and obtain the associated confidence bands empirically using 1000 sample paths for each forecast.

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   \includegraphics[width=0.8\linewidth]{simulated/12hr/1124.pdf}
   \includegraphics[width=0.8\linewidth]{simulated/24hr/19.pdf}  %569
   %\includegraphics[width=1\linewidth]{simulated/24hr/1099.pdf}
   %\includegraphics[width=0.9\linewidth]{6hr_forecast_CI_68.pdf}
\end{center}
   \caption{ Examples of simulated paths of wind power production for the next 12 and 24 hours.}
\label{simulated_24hr}
\end{figure}

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{confidence_intervals/24hr/31.pdf}
  \includegraphics[width=0.8\linewidth]{confidence_intervals/24hr/820.pdf}
  %\includegraphics[width=1\linewidth]{confidence_intervals/24hr/82.pdf}

\end{center}
  \caption{ Examples of confidence bands obtained for 24 hour forecasts. We can see that the model captures the fluctuations in the actual production with non-trivial and asymmetric confidence intervals.}
\label{fig:72hr}
\end{figure}

\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{confidence_intervals/12hr/75.pdf}
  \includegraphics[width=0.8\linewidth]{confidence_intervals/12hr/148.pdf}
  %\includegraphics[width=0.9\linewidth]{6hr_forecast_CI_68.pdf}
\end{center}
  \caption{ Examples of confidence bands obtained for the first 12 hours of the forecasts. Note that this forecasting company computes a new forecast every 6-9 hours for reliability.}
\label{fig:6hr}
\end{figure}

@@

\section*{Conclusions}
In this project, we have proposed a model based on parametric Stochastic Differential Equations and advanced inference to quantify uncertainties in wind power generation forecasts. It has the following advantages:
\begin{itemize}
\item  Represents and quantifies uncertainty in wind power forecasts accordingly to their real-world performance.
\item Ability to simulate paths of wind power production.
\item It is forecast technology agnostic.
\item  Takes into account physical constrains and the skew-symmetric nature of forecast error.
\item Provides a basis for decision making in the optimal dispatch of electric power.
\end{itemize}

@@

% \section*{References}
% \begin{flushleft}
% [1] {Elkantassi, S., Kalligiannaki, E., \& Tempone, R. (2017). Inference And Sensitivity In Stochastic Wind Power Forecast Models. Proceedings of the 2nd International Conference on Uncertainty Quantification in Computational Sciences and Engineering (UNCECOMP 2017). doi:10.7712/120217.5377.16899}
% \end{flushleft}
